---
title: "Nov 2024 - Wisconsin Election Results"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, echo = F)
library(dplyr)
library(tidyr)
library(sf)
library(leaflet)
library(gt)

df <- st_read(here::here("2024-nov/processed/matched-reporting-unit-results.geojson"))

df.state <- df |>
  st_drop_geometry() |>
  summarise(across(ends_with("24"), ~sum(.x, na.rm = T)),
            across(ends_with("20"), ~sum(.x, na.rm = T))) |>
  mutate(across(where(is.numeric), round),
         PREmarginv24 = PREDEM24 - PREREP24,
         PREmarginp24 = (PREDEM24/PRETOT24 - PREREP24/PRETOT24)*100,
         USSmarginv24 = USSDEM24 - USSREP24,
         USSmarginp24 = (USSDEM24/USSTOT24 - USSREP24/USSTOT24)*100,
         PREmarginv20 = PREDEM20 - PREREP20,
         PREmarginp20 = (PREDEM20/PRETOT20 - PREREP20/PRETOT20)*100)
```

This page shows the current status of our election night data collection. We are collecting, cleaning, and standardizing ward-level (reporting units, technically) results from counties across Wisconsin. Once a reporting unit is processed, it is added to the table and map below.

Please note, [these municipalities](https://elections.wi.gov/central-count-absentee-ballot-municipalities) use a central location to count their absentee ballots. Those results will be added to the appropriate ward totals for each voter. However, the initial ward results shown here may not yet include the central count totals.

***THIS WEBPAGE CURRENTLY INCLUDES ONLY TEMPLATE DATA--NOT ACTUAL ELECTION RESULTS.***

```{r}
df.state |>
  pivot_longer(cols = everything()) |>
  separate(name, into = c("office", "party","year"), sep = c(3, -2)) |>
  mutate(office = paste0(office, " 20", year)) |>
  select(-year) |>
  pivot_wider(names_from = party, values_from = value) |>
  gt(rowname_col = "office") |>
  fmt_number(columns = c(TOT, DEM, REP, marginv), decimals = 0) |>
  fmt_number(columns = marginp, decimals = 1, pattern = "{x} pt.") |>
  sub_missing() |>
  tab_style(style = cell_text(color = "red", weight = "bold"),
            locations = cells_body(columns = marginv,
                                   rows = marginv < 0)) |>
  tab_style(style = cell_text(color = "red", weight = "bold"),
            locations = cells_body(columns = marginp,
                                   rows = marginp < 0)) |>
  tab_style(style = cell_text(color = "blue", weight = "bold"),
            locations = cells_body(columns = marginv,
                                   rows = marginv > 0)) |>
  tab_style(style = cell_text(color = "blue", weight = "bold"),
            locations = cells_body(columns = marginp,
                                   rows = marginp > 0)) |>
  tab_header(title = "Election results in collected reporting units",
             subtitle = "see map for coverage area") |>
  tab_source_note("Ward boundaries change between elections. Results from old wards are distributed into the 2024 boundaries based on the number of residents in the overlapping sections.")
```


```{r}
df |>
  st_transform(crs = 4326) |>
  leaflet() |>
  addProviderTiles(provider = providers$CartoDB.Positron) |>
  addPolygons(label = ~reporting_unit)
```
